---
title: "simpleRegress"
author: "Connor Rice"
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

The prediction of overall length based on neck dimensions has a residual standard error of: 5.376 and an adjusted r-squared value of: 0.7476.

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp1 <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp1)
summary.lm(simp1)
```

The prediction of overall length based on age of the bear has a residual standard error of: 7.511 and an adjusted r-squared value of: 0.5073.

```{r}
simp2 <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp2)
summary.lm(simp2)
```

The prediction of overall length based on head length dimensions has a residual standard error of: 4.235 and an adjusted r-squared value of: 0.8434.

```{r}
simp3 <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp3)
summary.lm(simp3)
```

The prediction of overall length based on head width dimensions has a residual standard error of: 7.320 and an adjusted r-squared value of: 0.5321.

```{r}
simp4 <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp4)
summary.lm(simp4)
```

The prediction of overall length based on chest dimensions has a residual standard error of: 4.939 and an adjusted r-squared value of: 0.7869.

```{r}
simp5 <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp5)
summary.lm(simp5)
```

The prediction of overall length based on weight of the bear has a residual standard error of: 5.432 and an adjusted r-squared value of: 0.7423.

## Conclusions & Numerical Ranking

The ranked variables, from best to worst, based on residual standard error(RSE) and adjusted r-squared(ARS) value as predictors of overall length of a bear are:
  
  1. Head Length: RSE: 4.325 ARS: 0.8434
  2. Chest Measurement: RSE: 4.939 ARS: 0.7869
  3. Neck Measurement: RSE: 5.376 ARS: 0.7476
  4. Weight: RSE: 5.432 ARS: 0.7423
  5. Head Width: RSE: 7.32 ARS: 0.5321
  6. Age: RSE: 7.511 ARS: 0.5073
  
This ranking/trend is implied by both the adjusted r-squared value of the variable comparisons decreasing in correspondence with their ranking and the residual standard error increasing in correspondence as their ranking decreases. Head length is the best predictor variable of the overall length of the bear and age is the worst predictor variable.